---
title: "Script1"
author: "Fernanda Gonzalez-Saldias"
date: "2025-06-02"
output:
  pdf_document:
    number_sections: yes
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---
```{r, include=F,message=F}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
```

# Set working space
## Load libraries
```{r message=FALSE, warning=FALSE}

rm(list=ls())
library(vegan)
library(ggplot2)
library(dplyr)
library(readxl)
library(xlsx)
library(Hmisc)
library(ggcorrplot)
library(ggthemes)
library(corrplot)
library(vegan)
library(cowplot)
library(car)
library("pgirmess")
library(microbiome)
library (adespatial)

```


## Load morphological and molecular of relative abundances dataset
```{r message=FALSE, warning=FALSE,fig.align="center"}
Morpho=data.frame(read_xlsx("data\\Table S2.xlsx",sheet="Morphological",na=""))
head(Morpho)[,1:5]
COI=data.frame(read_xlsx("data\\Table S2.xlsx",sheet="COI",na=""))
head(COI)[,1:5]
X18s=data.frame(read_xlsx("data\\Table S2.xlsx",sheet="18S rRNA",na=""))
head(X18s)[,1:5]

#assigned rownames
rownames(Morpho)=(Morpho$Station)
rownames(COI)=(COI$Station)
rownames(X18s)=(X18s$Station)

#Select the numeric data
dato_morpho<-Morpho[4:386]
dato_coi<-COI[4:280]#277 otus
dato_18s<-X18s[4:169]#166 otus

#square root transformation of relative abundance data
datos.trans_morpho <- sqrt(dato_morpho)
datos.trans_coi <- sqrt(dato_coi)
datos.trans_18s<- sqrt(dato_18s)
```

# Procrustes analysis
## Distances calculation
```{r message=FALSE, warning=FALSE}
dist.BC_morpho <- vegdist(datos.trans_morpho)
dist.BC_coi <- vegdist(datos.trans_coi) 
dist.BC_18s <- vegdist(datos.trans_18s) 

#distances
out.mds_morpho <- cmdscale(dist.BC_morpho,eig=T,k=2)
out.mds_coi <- cmdscale(dist.BC_coi,eig=T,k=2)
out.mds_18s <- cmdscale(dist.BC_18s ,eig=T,k=2)
```


## Procrustes between morphological to molecular identification
```{r message=FALSE, warning=FALSE}
#Residual calculation
#COI
pro_morpho_coi<- procrustes(X = out.mds_morpho,
                            Y = out.mds_coi, symmetric = TRUE,choices = c(1,2))
Res.coi=data.frame(resid=residuals(pro_morpho_coi),
                   Station=names(residuals(pro_morpho_coi)),Method="COI")

#18S
pro_morpho_18s<- procrustes(X = out.mds_morpho, 
                            Y = out.mds_18s, symmetric = TRUE)
Res.18s=data.frame(resid=residuals(pro_morpho_18s),
                   Station=names(residuals(pro_morpho_18s)),Method="18S")

#Concatenated residual of procrustes dataframe
Res.all<-rbind(Res.coi,Res.18s)

#a merge is performed with the data matrix to add the location, station and pH.
Res.all=merge(Res.all,Morpho[,1:3],by="Station",sort = FALSE)

#Set method as a factor
Res.all$Method=as.factor(Res.all$Method)

#reorder factors
Res.all$Method=factor(Res.all$Method,c("COI","18S"))
head(Res.all)
```


## Create the residual figure
```{r fig.align="center", message=FALSE, warning=FALSE}
#Display panel figure 3C
Residuals=Res.all%>%
  ggplot(aes(Method,resid))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(width = 0.2,shape=21,size=2)+labs(x=" ",y="Residuals",title = "")+
  ylim(0,0.18)+
  theme_bw()+ theme(strip.text.x = element_text(size=12),
                    strip.background = element_blank(),
                    panel.grid.minor = element_blank(),
                    panel.grid.major = element_blank(), 
                    axis.text.x = element_text(size=12,color="black",
                                               angle=0,vjust = 0.5),
                    axis.text.y = element_text(size=8,color="black",
                                               angle=0,vjust = 0.5),
                    axis.title.x = element_text(color="black", size=10,
                                                margin = margin(t = 5)),
                    axis.title.y = element_text(color="black", size=10),
                    text = element_text(family="serif"))
```



```{r message=FALSE, warning=FALSE}
#Normality test
tapply(Res.all$resid,Res.all$Method,shapiro.test)
#Statistical test
wilcox.test(Res.all$resid~Res.all$Method)

```


## Procrustes calculation
```{r message=FALSE, warning=FALSE}
#Morphological v/s COI procrustes calculation
pro_morpho_coi<- procrustes(X = out.mds_morpho, 
                            Y = out.mds_coi, symmetric = TRUE,choices = c(1,2))
#stadistical test
protest(X = out.mds_morpho, 
        Y = out.mds_coi, scores = "sites", permutations = 999)

#Get data from pro_morfo_coi
tmp.pro<-rbind(pro_morpho_coi[["Yrot"]],pro_morpho_coi[["X"]])

pro.df<-data.frame(X=tmp.pro[,1],
                   Y=tmp.pro[,2],
                   Method=rep(c("Molecular","Morphological"),each=66),
                   Station=row.names(tmp.pro))

pro.df$Method<-as.factor(pro.df$Method)
pro.df$Method<-factor(pro.df$Method, c("Molecular","Morphological"))

pro.df<-merge(pro.df,COI[,1:3],by="Station",sort = FALSE)

#Display panel figure 3A
pro_morpho_coi=pro.df%>%
  ggplot(aes(X,Y,color=pH,shape=Method))+
  geom_line(aes(group=Station),color="black")+
  geom_point(size=4)+
  theme_bw()+labs(x="Dimension 1",y="Dimension 2",title = "Morphological vs. COI")+
  theme(strip.text.x = element_text(size=12),strip.background = element_blank(),
        panel.grid.minor = element_blank(),panel.grid.major = element_blank(),
        plot.title = element_text(size=12,hjust =0.5), 
        axis.title.x =element_text(size=10,color="black"),
        axis.title.y= element_text(size=10,color="black"),
        axis.text.x = element_text(size=8,color="black",angle=0,vjust = 0.5),
        axis.text.y = element_text(size=8,color="black",angle=0,vjust = 0.5),
        text = element_text(family="serif"))+
  scale_color_gradientn(colours = c("#F27F0C","yellow","#7ed348","#26B170","#01377d"))+
  theme(legend.position = "none")
```


```{r message=FALSE, warning=FALSE}
#Morphological v/s 18S procrustes calculation
pro_morpho_18s<- procrustes(X = out.mds_morpho, Y = out.mds_18s, symmetric = TRUE)
#Statistical test
protest(X = out.mds_morpho, Y = out.mds_18s, scores = "sites", permutations = 999)

#Get data from pro_morfo_18s
tmp.pro_2<-rbind(pro_morpho_18s[["Yrot"]],pro_morpho_18s[["X"]])

pro.df_2<-data.frame(X=tmp.pro_2[,1],
                     Y=tmp.pro_2[,2],
                     Method=rep(c("Molecular","Morphological"),each=66),
                     Station=row.names(tmp.pro_2))

pro.df_2$Method<-as.factor(pro.df_2$Method)
pro.df_2$Method<-factor(pro.df_2$Method, c("Molecular","Morphological"))

pro.df_2<-merge(pro.df_2,X18s[,1:3],by="Station",sort = FALSE)

#Display panel figure 3B
pro_morpho_18s=pro.df_2%>%
  ggplot(aes(X,Y,color=pH,shape=Method))+
  geom_line(aes(group=Station),color="black")+
  geom_point(size=4)+
  theme_bw()+labs(x="Dimension 1",y="Dimension 2",title = "Morphological vs. 18S rRNA")+
  theme(strip.text.x = element_text(size=12),strip.background = element_blank(),
        panel.grid.minor = element_blank(),panel.grid.major = element_blank(),
        plot.title = element_text(size=12,hjust =0.5), 
        axis.title.x =element_text(size=10,color="black"),
        axis.title.y= element_text(size=10,color="black"),
        axis.text.x = element_text(size=8,color="black",angle=0,vjust = 0.5),
        axis.text.y = element_text(size=8,color="black",angle=0,vjust = 0.5),
        text = element_text(family="serif"))+
  scale_color_gradientn(colours = c("#F27F0C","yellow","#7ed348","#26B170","#01377d"))+
  theme(legend.position = "bottom",legend.key.size = unit(0.5, 'cm'),
        legend.text = element_text(size=10),
        legend.title = element_text(size = 12))+
  guides(shape = guide_legend(title.position = "top",ncol=1))
legenda_18s_pro=get_plot_component(pro_morpho_18s, 'guide-box-bottom', return_all = TRUE)
pro_morpho_18s=pro_morpho_18s+theme(legend.position = "none")
```

## Display Figure 3
Figure 3: Comparison of MDS through Procrustes analysis between morphological vs. COI (A) and morphological vs. 18S (B). Residual values from comparing morphological identification and the two molecular identifications (18S and COI) (C).

```{r echo=FALSE, fig.align="center", message=FALSE, warning=FALSE}
#pdf(file="Figure 3.pdf",height = 15/2.54, width = 20/2.54)
plot_grid(plot_grid(pro_morpho_coi,Residuals,labels = c("A","C"),label_fontfamily = "serif"),
          plot_grid(pro_morpho_18s,legenda_18s_pro,labels = c("B"),label_fontfamily =  "serif"),nrow =2,
          rel_heights = c(0.9,0.9),rel_widths = c(0.8,0.1),axis = "l")
#dev.off()
```


# Lineal model
## Load metadata, alpha diversity calculation and Statistical analysis
```{r message=FALSE, warning=FALSE}
#Load metadata
metadata=data.frame(read_xlsx("data\\Table S1.xlsx",sheet="Metadata",na=""))

#alpha diversity calculation
metadata$H_morpho=vegan::diversity(t(t(Morpho[,4:386])))
metadata$S_morpho=vegan::specnumber(t(t(Morpho[,4:386])))
metadata$J_morpho=evenness((t(Morpho[,4:386])),'pielou')[,1]

metadata$H_COI=vegan::diversity(t(t(COI[,4:280])))
metadata$S_COI=vegan::specnumber(t(t(COI[,4:280])))
metadata$J_COI=evenness((t(COI[,4:280])),'pielou')[,1]

metadata$H_18s=vegan::diversity(t(t(X18s[,4:169])))
metadata$S_18s=vegan::specnumber(t(t(X18s[,4:169])))
metadata$J_18s=evenness((t(X18s[,4:169])),'pielou')[,1]
head(metadata)[,1:5]

morpho_index=metadata[,1:2]
morpho_index$Type<-"Morpho"
morpho_index$Richness=metadata[,25]
morpho_index$Shannon=metadata[,24]
morpho_index$Pielou=metadata[,26]

coi_index=metadata[,1:2]
coi_index$Type<-"COI"
coi_index$Richness=metadata[,28]
coi_index$Shannon=metadata[,27]
coi_index$Pielou=metadata[,29]

x18s_index=metadata[,1:2]
x18s_index$Type<-"18S"
x18s_index$Richness=metadata[,31]
x18s_index$Shannon=metadata[,30]
x18s_index$Pielou=metadata[,32]

analisis=rbind(morpho_index,coi_index,x18s_index)
head(analisis)[,1:5]

```


```{r message=FALSE, warning=FALSE,fig.align="center"}
#Richness
#Normality test
tapply(analisis$Richness,analisis$Type,shapiro.test)

#homoscedasticity test
bartlett.test(analisis$Richness~analisis$Type)

#kruskal Wallis test calculation
kruskal_S=kruskal.test(analisis$Richness~analisis$Type)
kruskal_S
#Post test to evaluate where are the significant differences
kruskalmc(analisis$Richness~analisis$Type,alpha=0.001)

```


```{r message=FALSE, warning=FALSE}
#Shannon index
#Normality test
tapply(analisis$Shannon,analisis$Type,shapiro.test)

#homoscedasticity test
bartlett.test(analisis$Shannon~analisis$Type)

#kruskal Wallis test calculation
kruskal_H=kruskal.test(analisis$Shannon~analisis$Type)
kruskal_H
#Post test to evaluate where are the significant differences
kruskalmc(analisis$Shannon~analisis$Type,alpha=0.01)

```


```{r message=FALSE, warning=FALSE}
#Pielou evenness
#Normality test
tapply(analisis$Pielou,analisis$Type,shapiro.test)

#homoscedasticity test
bartlett.test(analisis$Pielou~analisis$Type)

#kruskal Wallis test calculation
kruskal_J=kruskal.test(analisis$Pielou~analisis$Type)
kruskal_J
```


## To make a new data frame with pH mean, minimum and max for location and alpha index
```{r message=FALSE, warning=FALSE}
ind_ph=aggregate(pH~Location,data=metadata,mean)
ind_ph$pH.sd=aggregate(pH~Location,data=metadata,sd)[,2]
ind_ph$S_morpho=aggregate(S_morpho~Location,data=metadata,mean)[,2]
ind_ph$S_morpho.sd=aggregate(S_morpho~Location,data=metadata,sd)[,2]
ind_ph$S_COI=aggregate(S_COI~Location,data=metadata,mean)[,2]
ind_ph$S_COI.sd=aggregate(S_COI~Location,data=metadata,sd)[,2]
ind_ph$S_18S=aggregate(S_18s~Location,data=metadata,mean)[,2]
ind_ph$S_18S.sd=aggregate(S_18s~Location,data=metadata,sd)[,2]

ind_ph$H_morpho=aggregate(H_morpho~Location,data=metadata,mean)[,2]
ind_ph$H_morpho.sd=aggregate(H_morpho~Location,data=metadata,sd)[,2]
ind_ph$H_COI=aggregate(H_COI~Location,data=metadata,mean)[,2]
ind_ph$H_COI.sd=aggregate(H_COI~Location,data=metadata,sd)[,2]
ind_ph$H_18S=aggregate(H_18s~Location,data=metadata,mean)[,2]
ind_ph$H_18S.sd=aggregate(H_18s~Location,data=metadata,sd)[,2]

ind_ph$J_morpho=aggregate(J_morpho~Location,data=metadata,mean)[,2]
ind_ph$J_morpho.sd=aggregate(J_morpho~Location,data=metadata,sd)[,2]
ind_ph$J_COI=aggregate(J_COI~Location,data=metadata,mean)[,2]
ind_ph$J_COI.sd=aggregate(J_COI~Location,data=metadata,sd)[,2]
ind_ph$J_18S=aggregate(J_18s~Location,data=metadata,mean)[,2]
ind_ph$J_18S.sd=aggregate(J_18s~Location,data=metadata,sd)[,2]
head(ind_ph)
```


## Models and figures
### Richness
```{r fig.align="center", fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
#Morphological
lineal.poly=lm(S_morpho~poly(pH,2),data=ind_ph)
summary(lineal.poly)
lineal.lm=lm(S_morpho~pH,data=ind_ph)
summary(lineal.lm)
AIC(lineal.poly,lineal.lm)

#Display panel figure 4A
S_morpho=ind_ph%>%
  ggplot(aes(pH,S_morpho))+
  geom_point() + 
  geom_errorbar(aes(ymin = S_morpho-S_morpho.sd,ymax = S_morpho+S_morpho.sd )) + 
  geom_errorbarh(aes(xmin = pH-pH.sd,xmax = pH+pH.sd))+
  theme_bw()+
  theme(strip.text.x = element_text(size=10),strip.background = element_blank(),
        panel.grid.minor = element_blank(),panel.grid.major = element_blank(),
        axis.text.x =element_blank(),axis.title.x= element_blank(),
        axis.text.y = element_text(size=10, color="black"),
        axis.title = element_text(color="black",size=12, face="bold"),
        text = element_text(family="serif"))+
  labs(y="Richness",x="pH", title = "      Morphological") +xlim(4,8)+ylim(0,70)+
  geom_smooth(se=FALSE,method="lm", formula=y~poly(x,2))+
  annotate("text",x=5.8,y=11,
           label=bquote("y="~.(round(summary(lineal.poly)[[4]][3,1],3))~
                          "x"^2~"+"~.(round(summary(lineal.poly)[[4]][2,1],3))~
                          "x"~"+"~.(round(summary(lineal.poly)[[4]][1,1],3))),
           cex = 2.5,col="black")+
  annotate("text",x=5.8,y=5,
           label=bquote("R"^2~"="~.(round(summary(lineal.poly)[[8]],3))),
           cex = 2.5,col="black")

```

```{r fig.align="center", fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
#COI
lineal.poly=lm(S_COI~poly(pH,2),data=ind_ph)
summary(lineal.poly)
lineal.lm=lm(S_COI~pH,data=ind_ph)
summary(lineal.lm)
AIC(lineal.poly,lineal.lm)

#Display panel figure 4B
S_COI=ind_ph%>%
  ggplot(aes(pH,S_COI))+
  geom_point() + 
  geom_errorbar(aes(ymin = S_COI-S_COI.sd,ymax = S_COI+S_COI.sd )) + 
  geom_errorbarh(aes(xmin = pH-pH.sd,xmax = pH+pH.sd))+
  theme_bw()+
  theme(strip.text.x = element_text(size=10),strip.background = element_blank(),
        panel.grid.minor = element_blank(),panel.grid.major = element_blank(),
        axis.text.x =element_blank(),axis.title.x= element_blank(),
        axis.text.y =element_blank(),axis.title.y= element_blank(),
        text = element_text(family="serif"))+
  labs(y="",x="pH", title = "      COI") +xlim(4,8)+ylim(0,70)+
  geom_smooth(se=FALSE,method="lm", formula=y~poly(x,2))+
  annotate("text",x=5.6,y=60,
           label=bquote("y="~.(round(summary(lineal.poly)[[4]][3,1],3))~
                        "x"^2~"+"~.(round(summary(lineal.poly)[[4]][2,1],3))~
                        "x"~"+"~.(round(summary(lineal.poly)[[4]][1,1],3))),
           cex = 2.5,col="black")+
  annotate("text",x=5.6,y=55,
           label=bquote("R"^2~"="~.(round(summary(lineal.poly)[[8]],3))~"*"),
           cex = 2.5,col="black")

```


```{r fig.align="center", fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
#18S rRNA
lineal.poly=lm(S_18S~poly(pH,2),data=ind_ph)
summary(lineal.poly)
lineal.lm=lm(S_18S~pH,data=ind_ph)
summary(lineal.lm)
AIC(lineal.poly,lineal.lm)

#Display panel figure 4C
S_18s=ind_ph%>%
  ggplot(aes(pH,S_18S))+
  geom_point() + 
  geom_errorbar(aes(ymin = S_18S-S_18S.sd,ymax = S_18S+S_18S.sd )) + 
  geom_errorbarh(aes(xmin = pH-pH.sd,xmax = pH+pH.sd))+
  theme_bw()+
  theme(strip.text.x = element_text(size=10),strip.background = element_blank(),
        panel.grid.minor = element_blank(),panel.grid.major = element_blank(),
        axis.text.x =element_blank(),axis.title.x= element_blank(),
        axis.text.y =element_blank(),axis.title.y= element_blank(),
        text = element_text(family="serif"))+
  labs(y="",x="pH", title = "      18S rRNA") +xlim(4,8)+ylim(0,70)+
  geom_smooth(se=FALSE,method="lm", formula=y~poly(x,2))+
  annotate("text",x=5.5,y=60,
           label=bquote("y="~.(round(summary(lineal.poly)[[4]][3,1],3))~
                          "x"^2~"+"~.(round(summary(lineal.poly)[[4]][2,1],3))~
                          "x"~"+"~.(round(summary(lineal.poly)[[4]][1,1],3))),
           cex = 2.5,col="black")+
  annotate("text",x=5.5,y=55,
           label=bquote("R"^2~"="~.(round(summary(lineal.poly)[[8]],3))~"*"),
           cex =2.5,col="black")

```


### Shannon index
```{r fig.align="center", fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
#Morpholoical
lineal.poly=lm(H_morpho~poly(pH,2),data=ind_ph)
summary(lineal.poly)
lineal.lm=lm(H_morpho~pH,data=ind_ph)
summary(lineal.lm)
AIC(lineal.poly,lineal.lm)

#Display panel figure 4D
H_morpho=ind_ph%>%
  ggplot(aes(pH,H_morpho))+
  geom_point() + 
  geom_errorbar(aes(ymin = H_morpho-H_morpho.sd,ymax = H_morpho+H_morpho.sd )) + 
  geom_errorbarh(aes(xmin = pH-pH.sd,xmax = pH+pH.sd))+
  theme_bw()+
  theme(strip.text.x = element_text(size=10),strip.background = element_blank(),
        panel.grid.minor = element_blank(),panel.grid.major = element_blank(),
        axis.text.x =element_blank(),axis.title.x= element_blank(),
        axis.text.y = element_text(size=10, color="black"),
        axis.title = element_text(color="black",size=12, face="bold"),
        text = element_text(family="serif"))+
  labs(y="Shannon",x="pH", title = "") +xlim(4,8)+ylim(0,4)+
  geom_smooth(se=FALSE,method="lm", formula=y~poly(x,2))+
  annotate("text",x=5.6,y=1.0,
           label=bquote("y="~.(round(summary(lineal.poly)[[4]][3,1],3))~
                         "x"^2~"+"~.(round(summary(lineal.poly)[[4]][2,1],3))~
                         "x"~"+"~.(round(summary(lineal.poly)[[4]][1,1],3))),
           cex =2.5,col="black")+
  annotate("text",x=5.6,y=0.7,
           label=bquote("R"^2~"="~.(round(summary(lineal.poly)[[8]],3))),
           cex =2.5,col="black")

```


```{r fig.align="center", fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
#COI
lineal.poly=lm(H_COI~poly(pH,2),data=ind_ph)
summary(lineal.poly)
lineal.lm=lm(H_COI~pH,data=ind_ph)
summary(lineal.lm)
AIC(lineal.poly,lineal.lm)

#Display panel figure 4E
H_COI=ind_ph%>%
  ggplot(aes(pH,H_COI))+
  geom_point() + 
  geom_errorbar(aes(ymin = H_COI-H_COI.sd,ymax = H_COI+H_COI.sd )) + 
  geom_errorbarh(aes(xmin = pH-pH.sd,xmax = pH+pH.sd))+
  theme_bw()+
  theme(strip.text.x = element_text(size=10),strip.background = element_blank(),
        panel.grid.minor = element_blank(),panel.grid.major = element_blank(),
        axis.text.x =element_blank(),axis.title.x= element_blank(),
        axis.text.y =element_blank(),axis.title.y= element_blank(),
        text = element_text(family="serif"))+
  labs(y="",x="pH", title = "") +xlim(4,8)+ylim(0,4)+
  geom_smooth(se=FALSE,method="lm", formula=y~poly(x,2))+
  annotate("text",x=5.5,y=3.9,
           label=bquote("y="~.(round(summary(lineal.poly)[[4]][3,1],3))~
                        "x"^2~"+"~.(round(summary(lineal.poly)[[4]][2,1],3))~
                        "x"~"+"~.(round(summary(lineal.poly)[[4]][1,1],3))),
           cex =2.5,col="black")+
  annotate("text",x=5.5,y=3.6,
           label=bquote("R"^2~"="~.(round(summary(lineal.poly)[[8]],3))~"**"),
           cex =2.5,col="black")

```


```{r fig.align="center", fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
#18S rRNA
lineal.poly=lm(H_18S~poly(pH,2),data=ind_ph)
summary(lineal.poly)
lineal.lm=lm(H_18S~pH,data=ind_ph)
summary(lineal.lm)
AIC(lineal.poly,lineal.lm)

#Display panel figure 4F
H_18s=ind_ph%>%
  ggplot(aes(pH,H_18S))+
  geom_point() + 
  geom_errorbar(aes(ymin = H_18S-H_18S.sd,ymax = H_18S+H_18S.sd )) + 
  geom_errorbarh(aes(xmin = pH-pH.sd,xmax = pH+pH.sd))+
  theme_bw()+
  theme(strip.text.x = element_text(size=10),strip.background = element_blank(),
        panel.grid.minor = element_blank(),panel.grid.major = element_blank(),
        #axis.text.x = element_text(size=10, color="black",hjust=1),
        #axis.title.x = element_text(color="black",size=12,face="bold"),
        axis.text.x =element_blank(),axis.title.x= element_blank(),
        axis.text.y =element_blank(),axis.title.y= element_blank(),
        text = element_text(family="serif"))+
  labs(y="",x="pH", title = "") +xlim(4,8)+ylim(0,4)+
  geom_smooth(se=FALSE,method="lm", formula=y~poly(x,2))+
  annotate("text",x=5.5,y=3.8,
           label=bquote("y="~.(round(summary(lineal.poly)[[4]][3,1],3))~
                         "x"^2~"+"~.(round(summary(lineal.poly)[[4]][2,1],3))~
                         "x"~"+"~.(round(summary(lineal.poly)[[4]][1,1],3))),
           cex =2.5,col="black")+
  annotate("text",x=5.5,y=3.5,
           label=bquote("R"^2~"="~.(round(summary(lineal.poly)[[8]],3))~"*"),
           cex =2.5,col="black")

```


### Pielou evenness
```{r fig.align="center", fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
#Morphological
lineal.poly=lm(J_morpho~poly(pH,2),data=ind_ph)
summary(lineal.poly)
lineal.lm=lm(J_morpho~pH,data=ind_ph)
summary(lineal.lm)
AIC(lineal.poly,lineal.lm)

#Display panel figure 4G
J_morpho=ind_ph%>%
  ggplot(aes(pH,J_morpho))+
  geom_point() + 
  geom_errorbar(aes(ymin = J_morpho-J_morpho.sd,ymax = J_morpho+J_morpho.sd )) + 
  geom_errorbarh(aes(xmin = pH-pH.sd,xmax = pH+pH.sd))+
  theme_bw()+
  theme(strip.text.x = element_text(size=10),strip.background = element_blank(),
        panel.grid.minor = element_blank(),panel.grid.major = element_blank(),
        axis.text.x = element_text(size=10, color="black",hjust=1),
        axis.title.x = element_text(color="black",size=12,face="bold"),
        axis.title = element_text(color="black",size=12, face="bold"),
        text = element_text(family="serif"))+
  labs(y="Pielou evenness",x="pH", title = "") +xlim(4,8)+ylim(0,1)+
  geom_smooth(se=FALSE,method="lm", formula=y~poly(x,2))+
  annotate("text",x=5.6,y=0.32,
           label=bquote("y="~.(round(summary(lineal.poly)[[4]][3,1],3))~
                          "x"^2~"+"~.(round(summary(lineal.poly)[[4]][2,1],3))~
                          "x"~"+"~.(round(summary(lineal.poly)[[4]][1,1],3))),
           cex =2.5,col="black")+
  annotate("text",x=5.6,y=0.25,
           label=bquote("R"^2~"="~.(round(summary(lineal.poly)[[8]],3))),
           cex =2.5,col="black")

```


```{r fig.align="center", fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
#COI
lineal.poly=lm(J_COI~poly(pH,2),data=ind_ph)
summary(lineal.poly)
lineal.lm=lm(J_COI~pH,data=ind_ph)
summary(lineal.lm)
AIC(lineal.poly,lineal.lm)

#Display panel figure 4H
J_COI=ind_ph%>%
  ggplot(aes(pH,J_COI))+
  geom_point() + 
  geom_errorbar(aes(ymin = J_COI-J_COI.sd,ymax = J_COI+J_COI.sd )) + 
  geom_errorbarh(aes(xmin = pH-pH.sd,xmax = pH+pH.sd))+
  theme_bw()+
  theme(strip.text.x = element_text(size=10),strip.background = element_blank(),
        panel.grid.minor = element_blank(),panel.grid.major = element_blank(),
        axis.text.x = element_text(size=10, color="black",hjust=1),
        axis.title.x = element_text(color="black",size=12,face="bold"),
        axis.text.y =element_blank(),axis.title.y= element_blank(),
        axis.title = element_text(color="black",size=12, face="bold"),
        text = element_text(family="serif"))+
  labs(y="",x="pH", title = "") +xlim(4,8)+ylim(0,1)+
  geom_smooth(se=FALSE,method="lm", formula=y~poly(x,2))+
  annotate("text",x=5.8,y=0.32,
           label=bquote("y="~.(round(summary(lineal.poly)[[4]][3,1],3))~
                         "x"^2~"+"~.(round(summary(lineal.poly)[[4]][2,1],3))~
                         "x"~"+"~.(round(summary(lineal.poly)[[4]][1,1],3))),
           cex = 2.5,col="black")+
  annotate("text",x=5.8,y=0.25,
           label=bquote("R"^2~"="~.(round(summary(lineal.poly)[[8]],3))~"***"),
           cex =2.5,col="black")

```


```{r fig.align="center", fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
#18S rRNA
lineal.poly=lm(J_18S~poly(pH,2),data=ind_ph)
summary(lineal.poly)
lineal.lm=lm(J_18S~pH,data=ind_ph)
summary(lineal.lm)
AIC(lineal.poly,lineal.lm)

#Display panel figure 4I
J_18s=ind_ph%>%
  ggplot(aes(pH,J_18S))+
  geom_point() + 
  geom_errorbar(aes(ymin = J_18S-J_18S.sd,ymax = J_18S+J_18S.sd )) + 
  geom_errorbarh(aes(xmin = pH-pH.sd,xmax = pH+pH.sd))+
  theme_bw()+
  theme(strip.text.x = element_text(size=10),strip.background = element_blank(),
        panel.grid.minor = element_blank(),panel.grid.major = element_blank(),
        axis.text.x = element_text(size=10, color="black",hjust=1),
        axis.title.x = element_text(color="black",size=12,face="bold"),
        axis.text.y =element_blank(),axis.title.y= element_blank(),
        axis.title = element_text(color="black",size=12, face="bold"),
        text = element_text(family="serif"))+
  labs(y="",x="pH", title = "") +xlim(4,8)+ylim(0,1)+
  geom_smooth(se=FALSE,method="lm", formula=y~poly(x,2))+
  annotate("text",x=6.6,y=0.30,
           label=bquote("y="~.(round(summary(lineal.poly)[[4]][3,1],3))~
                         "x"^2~"+"~.(round(summary(lineal.poly)[[4]][2,1],3))~
                         "x"~"+"~.(round(summary(lineal.poly)[[4]][1,1],3))),
           cex = 2.5,col="black")+
  annotate("text",x=6.6,y=0.23,
           label=bquote("R"^2~"="~.(round(summary(lineal.poly)[[8]],3))~"***"),
           cex =2.5,col="black")

```


## Display Figure 4
Figure 4: Correlation between pH and alpha diversity variables (Richness, Shannon index, and Pielou) through morphological (A, D, G), COI (B, E, H), and 18S rRNA (C, F, I) datasets ordered by increasing pH. The dot is the average pH vs the average of the corresponding alpha diversity variable of each locality. Lines are the standard deviations of pH on the horizontal axis and the corresponding alpha on the vertical axis. The blue line is an adjusted polynomial model of order two.

```{r echo=FALSE, fig.align="center", message=FALSE, warning=FALSE}
#pdf(file="Figure 4.pdf",height = 16.9/2.54, width = 16.9/2.54)
plot_grid(plot_grid
          (S_morpho,
            H_morpho,
            J_morpho,
            align = "v",axis="l",ncol=1,rel_heights=c(0.8,0.8,0.9),
            labels = c("A","D","G"),label_size = 12,label_fontfamily ="serif"),
          plot_grid(S_COI,
                    H_COI,
                    J_COI,
                    align = "v",axis="l",ncol=1,rel_heights=c(0.8,0.8,0.9),
                    labels = c("B","E","H"),label_size = 12,label_fontfamily ="serif"),
          plot_grid(S_18s,
                    H_18s,
                    J_18s,
                    align = "v",ncol=1,
                    rel_heights=c(0.8,0.8,0.9),
                    labels = c("C","F","I"),label_size = 12,label_fontfamily ="serif"),
          ncol=3,rel_widths = c(0.92,0.75,0.75))
#dev.off()

```


# Correlation and log2 fold changes
## Log2 fold changes
```{r message=FALSE, warning=FALSE}
meta.diff.coi<-metadata[,1:2]
meta.diff.18s<-metadata[,1:2]

#Richnnes log2 fold changes
meta.diff.coi$L2FC<-log2(metadata$S_COI/metadata$S_morpho)
meta.diff.coi$Method<-"COI"
meta.diff.18s$L2FC<-log2(metadata$S_18s/metadata$S_morpho)
meta.diff.18s$Method<-"18S"

meta.diff.rich<-rbind(meta.diff.coi,meta.diff.18s)
head(meta.diff.rich)
```


```{r fig.align="center", fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
#Display panel figure 2B
plot.L2FC.rich<-meta.diff.rich%>%
  ggplot(aes(L2FC,Method,fill=Method))+
  geom_boxplot(outlier.shape = NA,alpha=0.4)+
  geom_jitter(aes(L2FC,Method,fill=Method),position=position_jitterdodge(0.4),
              shape=21,size=1)+
  scale_fill_manual(values = c("#6600CC","#FF9900"))+geom_vline(xintercept = 0,
                                                                linetype="dashed")+
  labs(x="",y="Richness")+xlim(-3.5,2)+
  theme_bw()+theme(strip.text.x = element_text(size=10),
                   strip.background = element_blank(),
                   panel.grid.minor = element_blank(),
                   panel.grid.major = element_blank(),
                   axis.text.x =element_blank(),
                   axis.title.x= element_blank(),
                   axis.text.y = element_text(size=8,color="black",
                                              angle=0,vjust = 0.5),
                   axis.title.y = element_text(color="black", size=10),
                   text = element_text(family="serif"))+
  theme(legend.position = "right",legend.key.size = unit(0.5, 'cm'),
        legend.text = element_text(size=8))
legenda_rich=get_legend(plot.L2FC.rich)
plot.L2FC.rich=plot.L2FC.rich+theme(legend.position = "none")

#Normality test
shapiro.test(meta.diff.rich$L2FC[meta.diff.rich$Method=="COI"])
shapiro.test(meta.diff.rich$L2FC[meta.diff.rich$Method=="18S"])
#t-student
t.test(meta.diff.rich$L2FC[meta.diff.rich$Method=="COI"],mu=0)
t.test(meta.diff.rich$L2FC[meta.diff.rich$Method=="18S"],mu=0)
```


```{r message=FALSE, warning=FALSE}
#Shannon index log2 fold changes
meta.diff.sha.coi<-metadata[,1:2]
meta.diff.sha.18s<-metadata[,1:2]

meta.diff.sha.coi$L2FC<-log2(metadata$H_COI/metadata$H_morpho)
meta.diff.sha.coi$Method<-"COI"
meta.diff.sha.18s$L2FC<-log2(metadata$H_18s/metadata$H_morpho)
meta.diff.sha.18s$Method<-"18S"

meta.diff.shannon<-rbind(meta.diff.sha.coi,meta.diff.sha.18s)
head(meta.diff.shannon)
```


```{r fig.align="center", fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
#Display panel figure 2D
plot.L2FC.shannon<-meta.diff.shannon%>%
  ggplot(aes(L2FC,Method,fill=Method))+
  geom_boxplot(outlier.shape = NA,alpha=0.4)+
  geom_jitter(aes(L2FC,Method,fill=Method),position=position_jitterdodge(0.4),
              shape=21,size=1)+
  scale_fill_manual(values = c("#6600CC","#FF9900"))+geom_vline(xintercept = 0,
                                                                linetype="dashed")+
  labs(x="",y="Shannon")+xlim(-3.5,2)+
  theme_bw()+theme(strip.text.x = element_text(size=10),
                   strip.background = element_blank(),
                   panel.grid.minor = element_blank(),
                   panel.grid.major = element_blank(),
                   axis.text.x =element_blank(),
                   axis.title.x= element_blank(),
                   axis.text.y = element_text(size=8,color="black",
                                              angle=0,vjust = 0.5),
                   axis.title.y = element_text(color="black", size=10),
                   text = element_text(family="serif"))+
  theme(legend.position = "right",legend.key.size = unit(0.5, 'cm'),
        legend.text = element_text(size=8))
legenda_shannon=get_legend(plot.L2FC.shannon)
plot.L2FC.shannon=plot.L2FC.shannon+theme(legend.position = "none")

#Normality test
shapiro.test(meta.diff.shannon$L2FC[meta.diff.shannon$Method=="COI"])
shapiro.test(meta.diff.shannon$L2FC[meta.diff.shannon$Method=="18S"])
#Statistical test
wilcox.test(meta.diff.shannon$L2FC[meta.diff.shannon$Method=="COI"],mu=0)
wilcox.test(meta.diff.shannon$L2FC[meta.diff.shannon$Method=="18S"],mu=0)
```


```{r message=FALSE, warning=FALSE}
#Pielou evenness log2 fold changes
meta.diff.pie.coi<-metadata[,1:2]
meta.diff.pie.18s<-metadata[,1:2]

meta.diff.pie.coi$L2FC<-log2(metadata$J_COI/metadata$J_morpho)
meta.diff.pie.coi$Method<-"COI"
meta.diff.pie.18s$L2FC<-log2(metadata$J_18s/metadata$J_morpho)
meta.diff.pie.18s$Method<-"18S"

meta.diff.pielou<-rbind(meta.diff.pie.coi,meta.diff.pie.18s)
head(meta.diff.pielou)
```


```{r fig.align="center", fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
#Display panel figure 2F
plot.L2FC.pielou<-meta.diff.pielou%>%
  ggplot(aes(L2FC,Method, fill=Method))+
  geom_boxplot(outlier.shape = NA,alpha=0.4)+
  geom_jitter(aes(L2FC,Method,fill=Method),position=position_jitterdodge(0.4),
              shape=21,size=1)+
  scale_fill_manual(values = c("#6600CC","#FF9900"))+geom_vline(xintercept = 0,
                                                                linetype="dashed")+
  labs(x=expression("Log" [2] * " fold change"),y="Pielou evenness")+xlim(-3.5,2)+
  theme_bw()+theme(strip.text.x = element_text(size=10),
                   strip.background = element_blank(),
                   panel.grid.minor = element_blank(),
                   panel.grid.major = element_blank(),
                   axis.text.x = element_text(size=8,color="black",
                                              angle=0,vjust = 0.5),
                   axis.text.y = element_text(size=8,color="black",
                                              angle=0,vjust = 0.5),
                   axis.title.x = element_text(color="black", size=10,
                                               margin = margin(t = 0,l=0)),
                   axis.title.y = element_text(color="black", size=10),
                   text = element_text(family="serif"))+ 
  guides(fill=guide_legend(ncol=2))+
  theme(legend.position="none",legend.key.size = unit(0.4, 'cm'),
        legend.text = element_text(size=4),legend.title = element_text(size=6))
legenda_pielou=get_legend(plot.L2FC.pielou)

#Normality test
shapiro.test(meta.diff.pielou$L2FC[meta.diff.pielou$Method=="COI"])
shapiro.test(meta.diff.pielou$L2FC[meta.diff.pielou$Method=="18S"])
#Statistical test
wilcox.test(meta.diff.pielou$L2FC[meta.diff.pielou$Method=="COI"],mu=0)
wilcox.test(meta.diff.pielou$L2FC[meta.diff.pielou$Method=="18S"],mu=0)

```


## Correlation analysis
```{r fig.align="center", fig.height=3, fig.width=3, message=FALSE, warning=FALSE}
#Richness
corr_rich<-rcorr(as.matrix(metadata[,c(25,28,31)]),type="spearman")
corr_rich
#significance
round(as.dist(corr_rich$P),4)

#Display panel figure 2A
richness_corr=ggcorrplot(corr_rich$r,type = "upper",
                         outline.col = "white",lab=TRUE,insig = "blank",lab_size = 3)+
  theme_bw()+
  labs(x="",y="")+
  theme_tufte(base_size = 7)+
  theme(strip.text.x = element_text(size=12),strip.background = element_blank(),
        panel.grid.minor = element_blank(),panel.grid.major = element_blank(),
        axis.title.x =element_blank(),axis.title.y= element_blank(),
        axis.text.x = element_text(size=8,color="black",angle=0,vjust = 0.5),
        axis.text.y = element_text(size=8,color="black",angle=0,vjust = 0.5),
        text = element_text(family="serif"))+
  theme(legend.position = "none",legend.key.size = unit(0.5, 'cm'),
        legend.text = element_text(size=8))

```


```{r fig.align="center", fig.height=3, fig.width=3, message=FALSE, warning=FALSE}
#Shannon index
corr_sha<-rcorr(as.matrix(metadata[,c(24,27,30)]),type="spearman")
corr_sha
#significance
round(as.dist(corr_sha$P),4)

#Display panel figure 2C
shannon_corr=ggcorrplot(corr_sha$r,type = "upper",
                        outline.col = "white",lab=TRUE,insig = "blank",lab_size = 3)+
  labs(x="",y="")+
  theme_bw()+
  theme_tufte(base_size = 7)+
  theme(strip.text.x = element_text(size=12),strip.background = element_blank(),
        panel.grid.minor = element_blank(),panel.grid.major = element_blank(),
        axis.title.x =element_blank(),axis.title.y= element_blank(),
        axis.text.x = element_text(size=8,color="black",angle=0,vjust = 0.5),
        axis.text.y = element_text(size=8,color="black",angle=0,vjust = 0.5),
        text = element_text(family="serif"))+
  theme(legend.position = "none",legend.key.size = unit(0.5, 'cm'),
        legend.text = element_text(size=8))

```


```{r fig.align="center", fig.height=3, fig.width=3, message=FALSE, warning=FALSE}
#Pielou evenness
corr_pie<-rcorr(as.matrix(metadata[,c(26,29,32)]),type="spearman")
corr_pie
#Significance
round(as.dist(corr_pie$P),4)

#Display panel figure 2E
pielou_corr=ggcorrplot(corr_pie$r,type = "upper",
                       outline.col = "white",lab=TRUE,insig = "blank",lab_size=3)+
  labs(x="",y="")+
  theme_bw()+
  theme_tufte(base_size = 7)+
  theme(strip.text.x = element_text(size=12),strip.background = element_blank(),
        panel.grid.minor = element_blank(),panel.grid.major = element_blank(),
        axis.title.x =element_blank(),axis.title.y= element_blank(),
        axis.text.x = element_text(size=8,color="black",angle=0,vjust = 0.5),
        axis.text.y = element_text(size=8,color="black",angle=0,vjust = 0.5),
        text = element_text(family="serif"))+
  theme(legend.position = "bottom",legend.key.size = unit(0.4, 'cm'),
        legend.text = element_text(size=8),
        legend.title = element_text(size = 8))
legenda_pielou_corr=get_plot_component(pielou_corr, 'guide-box-bottom',
                                       return_all = TRUE)
pielou_corr=pielou_corr+theme(legend.position = "none")

```


## Display Figure 2
Figure 2: Correlation between molecular and morphological identification in richness (A), Shannon index (C), and Pielou evenness (E). Comparison of alpha diversity indices in the three approaches to diatom communities of high mountain mires in richness (B), Shannon index (D), and Pielou evenness (F), molecular with morphological identification.

```{r echo=FALSE, fig.align="center", message=FALSE, warning=FALSE}
#pdf(file="Figure 2.pdf",height = 12/2.54, width = 13/2.54)
plot_grid(plot_grid(richness_corr,shannon_corr,pielou_corr,legenda_pielou_corr,
                    align = "v",ncol =1,rel_heights=c(0.45,0.45,0.45,0.1),labels = c("A","C","E",""),
                    label_fontfamily = "serif",
                    hjust=c(-0.7,-0.7,-0.7,-0.7),vjust =c(1,1,1,1)),
          plot_grid(plot.L2FC.rich,plot.L2FC.shannon,plot.L2FC.pielou,NULL,
                    rel_heights=c(0.45,0.45,0.5,0.1),ncol=1,labels = c("B","D","F",""),
                    label_fontfamily = "serif",
                    hjust=c(-0.1,-0.1,-0.1,-0.1),vjust =c(1,1,1,1)),
          align = "v",ncol =2,rel_widths = c(0.45,0.5))
#dev.off()

```

# dbRDA
```{r message=FALSE, warning=FALSE,fig.align="center"}
Metadata=data.frame(read_xlsx("data\\Table S1.xlsx",sheet="Metadata",na=""))

x<-Metadata[3:23]
head(x)

#data transformation
datos.trans_morpho <- sqrt(dato_morpho)
datos.trans_coi <- sqrt(dato_coi)
datos.trans_18s<- sqrt(dato_18s)
```

## dbRDA Morphology
```{r message=FALSE, warning=FALSE,fig.align="center"}
analisis_morpho <- capscale(datos.trans_morpho ~ WTD+GDD+Rad+
                        Sphagnum+Brown_mosses+Acrocarp_mosses+Liverworts+
                        Bryophytes+Vascular_plants+
                       pH+Prec+K+Mg+Mn+Na+P+S+Si, Metadata,
                     dist="bray")

anova (analisis_morpho)  # 0.001 *** - it is significant
adjR2.tbrda <- RsquareAdj (analisis_morpho)$adj.r.squared 


#Forward selection
sel.fs <- forward.sel (Y = datos.trans_morpho , X=x, adjR2thresh = adjR2.tbrda) 
sel.fs

```

## dbRDA COI
```{r message=FALSE, warning=FALSE,fig.align="center"}
analisis_coi <- capscale(datos.trans_coi ~ WTD+GDD+Rad+
                        Sphagnum+Brown_mosses+Acrocarp_mosses+Liverworts+
                        Bryophytes+Vascular_plants+
                       pH+Prec+K+Mg+Mn+Na+P+S+Si, Metadata,
                     dist="bray")

anova (analisis_coi)  # 0.001 *** - it is significant
adjR2.tbrda <- RsquareAdj (analisis_coi)$adj.r.squared 

#Forward selection
sel.fs <- forward.sel (Y = datos.trans_coi , X=x, adjR2thresh = adjR2.tbrda) 
sel.fs

```

## dbRDA 18S
```{r message=FALSE, warning=FALSE,fig.align="center"}
analisis_18S <- capscale(datos.trans_18s ~ WTD+GDD+Rad+
                        Sphagnum+Brown_mosses+Acrocarp_mosses+Liverworts+
                        Bryophytes+Vascular_plants+
                       pH+Prec+K+Mg+Mn+Na+P+S+Si, Metadata,
                     dist="bray")

anova (analisis_18S)  # 0.001 *** - it is significant
adjR2.tbrda <- RsquareAdj (analisis_18S)$adj.r.squared 

#Forward selection
sel.fs <- forward.sel (Y = datos.trans_18s , X=x, adjR2thresh = adjR2.tbrda) 
sel.fs

```
